Simulating High-Dimensional Multivariate Data

### 2020-11-05 040-monte-carlo-evaluation.Rmd
2nd draft. polish bivariate plot

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        ## AGS. 17 Oct 2020. cor_bounds failed.
        ## I'll create a minimal working example and open an issue
        ## tmp_bounds <- cor_bounds(margins, type = type)
        ## cor_lo <- tmp_bounds$lower[1,2] + eps
        ## cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_lo <- -0.75
        cor_hi <- 0.95
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

```{r ch040-combineBiSims}
## compare relative differences in MSE for each dependency, compared to Pearson
allDat <- NULL
dat <- readRDS("./results/biNorm_sims.rds") %>% select( margins, type, N, rho, rho_hat  ) 
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biGamma_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biNB_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
allDat$margins <- factor(allDat$margins, levels = c( 'norm', 'gamma', 'nbinom') )
allDat$type <- factor(allDat$type, levels = c( 'pearson', 'kendall', 'spearman' ) )
allDat$N <- factor(allDat$N, levels = c( 1000, 10000, 100000) )
```

```{r ch040-bPlot, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.asp = 1.30, fig.weight = 4, fig.align='center', fig.cap = "Bivariate simulations match specified correlations. The horizontal axis plots the specified target correlations across the entire range of possible correlations for each bivariate margin. Normal margins are plotted in dark grey, gamma in medium grey, and negative binomial in light grey. As the number of simulated vectors $B$ increases from left to right, the variation in estimated correlations (vertical axis) decreases. The dashed line indicates equality between the specified and estimated correlations. Only the normal margins match the Pearson correlations (top row) exactly across the entire range of correlations. The two non-normal bivariate random vectors experience attentuation (downward bias) for extreme Pearson correlations (due to limitations in our algorithm) and more restriction (due to the Frechet bounds). The rank-based correlations (bottom two rows) are matched exactly for all margins and obtain the full range of possible dependencies."}
# https://www.datanovia.com/en/blog/how-to-change-ggplot-facet-labels/
# New facet label names
repsLabs <- paste0("B=", c("1,000", "10,000", "100,000") )
names(repsLabs) <- c(1000, 10000, 100000)
typeLabs <- c( 'Pearson', 'Kendall', 'Spearman' ) 
names(typeLabs) <- c( 'pearson', 'kendall', 'spearman' )

# Set colors
## RColorBrewer::display.brewer.all()
numColors <- 4
numGroups <- length(levels(allDat$margins))
myColors <- rev( RColorBrewer::brewer.pal(n = numColors, name = 'Greys')[ ((numColors - numGroups) + 1): numColors  ] )

>>>>

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        ## AGS. 17 Oct 2020. cor_bounds failed.
        ## I'll create a minimal working example and open an issue
        ## work around for now
        if ( type == 'pearson' ) {
            tmp_bounds <- cor_bounds(margins, type = type)
            cor_lo <- tmp_bounds$lower[1,2] + eps
            cor_hi <- tmp_bounds$upper[1,2] - eps
        } else {
            cor_lo <- -1 + eps
            cor_hi <- 1 - eps
        }
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

```{r ch040-combineBiSims, cache=F}
## compare relative differences in MSE for each dependency, compared to Pearson
allDat <- NULL
dat <- readRDS("./results/biNorm_sims.rds") %>% select( margins, type, N, rho, rho_hat  ) 
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biGamma_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biNB_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
allDat$margins <- factor(allDat$margins, levels = c( 'norm', 'gamma', 'nbinom') )
allDat$type <- factor(allDat$type, levels = c( 'pearson', 'kendall', 'spearman' ) )
allDat$N <- factor(allDat$N, levels = c( 1000, 10000, 100000) )
```

```{r ch040-bPlot, cache=F, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.asp = 1.30, fig.weight = 4, fig.align='center', fig.cap = "Bivariate simulations match specified correlations. The horizontal axis plots the specified target correlations across the entire range of possible correlations for each bivariate margin. Normal margins are plotted in dark grey, gamma in medium grey, and negative binomial in light grey. As the number of simulated vectors $B$ increases from left to right, the variation in estimated correlations (vertical axis) decreases. The dashed line indicates equality between the specified and estimated correlations. Only the normal margins match the Pearson correlations (top row) exactly across the entire range of correlations. The two non-normal bivariate random vectors experience attentuation (downward bias) for extreme Pearson correlations (due to limitations in our algorithm) and more restriction (due to the Frechet bounds). The rank-based correlations (bottom two rows) are matched exactly for all margins and obtain the full range of possible dependencies."}
# https://www.datanovia.com/en/blog/how-to-change-ggplot-facet-labels/
# New facet label names
repsLabs <- paste0("B=", c("1,000", "10,000", "100,000") )
names(repsLabs) <- c(1000, 10000, 100000)
typeLabs <- c( 'Pearson', 'Kendall', 'Spearman' ) 
names(typeLabs) <- c( 'pearson', 'kendall', 'spearman' )

# Set colors
## RColorBrewer::display.brewer.all()
numColors <- 4
numGroups <- length(levels(allDat$margins))
## myColors <- rev( RColorBrewer::brewer.pal(n = numColors, name = 'Greys')[ ((numColors - numGroups) + 1): numColors  ] )
myColors <- rev( RColorBrewer::brewer.pal(n = numColors, name = 'Blues')[ ((numColors - numGroups) + 1): numColors  ] )
